---
title: "DataFirstLook"
author: '500586901'
date: "28/03/2022"
output: html_document
---

```{r}
library(GEOquery) 
library(R.utils)
library(reshape2)
library(ggplot2)
library(limma)
library(dplyr)
library(devtools)
library(prideR)
```


```{r}
gse129166 <- getGEO("GSE129166", GSEMatrix = T, getGPL = F)
gse129166 <- gse129166$GSE129166_series_matrix.txt.gz


head(pData(gse129166))
fData(gse129166)
head(exprs(gse129166))


```

```{r}
gse13440 <- getGEO("GSE13440", GSEMatrix = T, getGPL = F)
gse13440 <- gse13440$GSE13440_series_matrix.txt.gz


head(pData(gse13440))
head(exprs(gse13440))

transpose_gse13440 <- t(exprs(gse13440)[2000:2100,])
View(transpose_gse13440)
boxplot(transpose_gse13440)
```


```{r}
gse129166 <- getGEO("GSE129166", GSEMatrix = T, getGPL = F)
gse129166 <- gse129166$GSE129166_series_matrix.txt.gz


head(pData(gse129166))
head(exprs(gse129166))

transpose_gse129166 <- t(exprs(gse129166)[1001:2000,])
boxplot(transpose_gse129166)

head(Meta(gse129166))
```

## The datasets I am checking

### GSE34437 biopsy
kinda weird as it has multiple characteristics - not just reject or stable, but reject, stable, baseline living donor, no significant abnormalities
66 people so not bad
Gene expression data is given as probes (e.g. 1007_s_at or 1255_g_at which may need some form of webscraping if we want to integrate this with actual gene names)
```{r}
gse34437 <- getGEO("GSE34437", GSEMatrix = T, getGPL = F)
gse34437 <- gse34437$GSE34437_series_matrix.txt.gz

pData(gse34437)
head(exprs(gse34437))
```

### GSE75693 biopsy
easy to clean the conditions but some weird ones - 'Absence of chronic allograft nephropathy' or 'BVKN'
79 obvs so not bad
Gene expression data is given as probes (e.g. 1007_s_at or 1255_g_at which may need some form of webscraping if we want to integrate this with actual gene names)
 
```{r}
gse75693 <- getGEO("GSE75693", GSEMatrix = T, getGPL = F)
gse75693 <- gse75693$GSE75693_series_matrix.txt.gz

pData(gse75693)
head(exprs(gse75693))
```

### GSE109564 biopsy
Doesn't work + also only 7 obs
```{r}
gse109564 <- getGEO("GSE109564", GSEMatrix = T, getGPL = F)
gse109564 <- gse109564$GSE109564_series_matrix.txt.gz

pData(gse109564)
head(exprs(gse109564))
```

### GSE112927 blood
Lots of observations but expression data is weird
```{r}
gse112927 <- getGEO("GSE112927", GSEMatrix = T, getGPL = F)
gse112927 <- gse112927$GSE112927_series_matrix.txt.gz

pData(gse112927)
head(exprs(gse112927))
fData(gse112927)
```

### GSE1563 blood
Good data set with 31 obs that looks consistent with the others in terms of gene markers
```{r}
gse1563 <- getGEO("GSE1563", GSEMatrix = T, getGPL = F)
gse1563 <- gse1563$GSE1563_series_matrix.txt.gz

pData(gse1563)
head(exprs(gse1563))
fData(gse1563)

transpose_gse1563 <- t(exprs(gse1563)[1001:1100,])
boxplot(transpose_gse1563, medcol = "red")
```

### GSE20300
only 24 samples but clear labelling of stable and unstable as well as same gene markers used
```{r}
gse20300 <- getGEO("GSE20300", GSEMatrix = T, getGPL = F)
gse20300 <- gse20300$GSE20300_series_matrix.txt.gz

pData(gse20300)
head(exprs(gse20300))
fData(gse20300)
```

### GSE22707
24 samples. Should eb fine to extract reject, stable based on which function
Same structure of genes as prev datasets
```{r}
gse22707 <- getGEO("GSE22707", GSEMatrix = T, getGPL = F)
gse22707 <- gse22707$GSE22707_series_matrix.txt.gz

pData(gse22707)
head(exprs(gse22707))
fData(gse22707)

transpose_gse22707 <- t(exprs(gse22707)[200:300,])
boxplot(transpose_gse22707, medcol="red")
```


### GSE98320
Doesn't have a reject or stable column
```{r}
gse98320 <- getGEO("GSE98320", GSEMatrix = T, getGPL = F)
gse98320 <- gse98320$GSE98320_series_matrix.txt.gz

pData(gse98320)
head(exprs(gse98320))
fData(gse98320)
```

### 