---
title: "Liam Research 1"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(class)
library(cvTools)
library(randomForest)
library(GEOquery) 
library(R.utils)
library(reshape2)
library(limma)
library(dplyr)
library(e1071)
library(DT)
library(viridis)
library(plotly)
library(scales)
library(CPOP)
library(matrixStats)
library(visNetwork)
library(sva)
```

## Mukund CPOP analysis 
```{r}
#CPOP data
GSE46474 = getGEO("GSE46474")[[1]]
GSE36059 = getGEO("GSE36059")[[1]]
GSE48581 = getGEO("GSE48581")[[1]]


# The below is what the user uploads
GSE129166 = getGEO("GSE129166")[[1]]

GSKhang = getGEO(filename="../Liam/Data/GSE129166_series_matrix.txt.gz")
slotNames(GSKhang)
# The below is not important
#GSE15296 = getGEO("GSE15296")[[1]]
```
# Create error function for user upload data

```{r}
## Khang's work 
u <- read_csv('../Liam/Data/userUploadData.csv')

set.seed(3888)

cname <- tail(colnames(u),-1)

amountOFErrors <- 10000

errors <- c(NA,"NA", "na","Na", "nA", "-","rubbish","brah", "/", "", "  ", "    ", "", "-99", -99,0)

for (val in cname){
  curVec <- u[[val]]
  ind <-  which(curVec %in% sample(curVec,amountOFErrors))
  u[[val]][ind] <- sample(errors,1)

}

write.csv(u,"Path to export the DataFrame\\File Name.csv", row.names = TRUE)
class(u)

```


```{r}
exp_GSE129166 = as.data.frame(exprs(GSE129166))
# 0 is stable, 1 is rejection
myP <- pData(GSE129166)
myP$diagnosis = ifelse(myP$characteristics_ch1.1 == "tcmr (no: 0_borderline:1_TCMR:2): 0" & myP$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0" , 0, 1)
myP["GSM3701701", "diagnosis"]

exp_GSE129166

output <- rbind(exp_GSE129166, t(cbind(myP[0], diagnosis = myP$diagnosis)))


# write.csv(output,"/Users/khangnguyen/Desktop/userUploadData.csv", row.names = TRUE)



```


```{r}


## keeping only the 100 most variable genes in my data frame 
exp_GSE36059 = (exprs(GSE36059))
Variance = rowVars(as.matrix(exp_GSE36059)) #
Variance = as.data.frame(Variance)

exp_GSE36059 = as.data.frame(exp_GSE36059)
class(exp_GSE36059)
rowVars(as.matrix(exp_GSE36059))

exp_GSE36059 = cbind(exp_GSE36059, variance = Variance)
exp_GSE36059 = slice_max(exp_GSE36059, order_by = Variance, n = 2000)
exp_GSE36059 = subset(exp_GSE36059, select = -c(Variance))
row_names_exp_GSE36059 = rownames(exp_GSE36059)

typeof(exp_GSE36059)

rownames(exp_GSE36059)




# write.csv(exp_GSE129166,"/Users/khangnguyen/Desktop/userUploadData.csv", row.names = TRUE)

exp_GSE46474 = (exprs(GSE46474))
Variance = rowVars(as.matrix(exp_GSE46474))
Variance = as.data.frame(Variance)
exp_GSE46474 = as.data.frame(exp_GSE46474)
exp_GSE46474 = cbind(exp_GSE46474, variance = Variance)
exp_GSE46474 = slice_max(exp_GSE46474, order_by = Variance, n = 2000)
exp_GSE46474 = subset(exp_GSE46474, select = -c(Variance))
row_names_exp_GSE46474 = rownames(exp_GSE46474)


exp_GSE48581 = (exprs(GSE48581))
Variance = rowVars(as.matrix(exp_GSE48581))
Variance = as.data.frame(Variance)
exp_GSE48581 = as.data.frame(exp_GSE48581)
exp_GSE48581 = cbind(exp_GSE48581, variance = Variance)
exp_GSE48581 = slice_max(exp_GSE48581, order_by = Variance, n = 2000)
exp_GSE48581 = subset(exp_GSE48581, select = -c(Variance))
row_names_exp_GSE48581 = rownames(exp_GSE48581)


intersection = intersect(row_names_exp_GSE36059, row_names_exp_GSE46474)
intersection = intersect(intersection, row_names_exp_GSE48581)

exp_GSE36059 = as.data.frame(t(as.matrix(exp_GSE36059)))
exp_GSE36059 = subset(exp_GSE36059, select = c(intersection))

exp_GSE46474 = as.data.frame(t(as.matrix(exp_GSE46474)))
exp_GSE46474 = subset(exp_GSE46474, select = c(intersection))

exp_GSE48581 = as.data.frame(t(as.matrix(exp_GSE48581)))
exp_GSE48581 = subset(exp_GSE48581, select = c(intersection))

GSE36059_id <- data.frame("Dataset" = rep("GSE36059",nrow(exp_GSE36059)))
GSE46474_id <- data.frame("Dataset" = rep("GSE46474",nrow(exp_GSE46474)))
GSE48581_id <- data.frame("Dataset" = rep("GSE48581",nrow(exp_GSE48581)))

z1 = exp_GSE36059 %>% as.matrix()
z3 = exp_GSE46474 %>% as.matrix()
z2 = exp_GSE48581 %>% as.matrix()

## arcsine transformation

exp_GSE36059_arc <- exp_GSE36059
exp_GSE36059_arc = exp_GSE36059_arc / max(exp_GSE36059_arc)
exp_GSE36059_arc = asin(sqrt(exp_GSE36059_arc))

exp_GSE46474_arc <- exp_GSE46474
exp_GSE46474_arc = exp_GSE46474_arc / max(exp_GSE46474_arc)
exp_GSE46474_arc = asin(sqrt(exp_GSE46474_arc))

exp_GSE48581_arc <- exp_GSE48581
exp_GSE48581_arc = exp_GSE48581_arc / max(exp_GSE48581_arc)
exp_GSE48581_arc = asin(sqrt(exp_GSE48581_arc))

z1_pairwise = pairwise_col_diff(z1) %>% as.matrix()
z2_pairwise = pairwise_col_diff(z2) %>% as.matrix()
z3_pairwise = pairwise_col_diff(z3) %>% as.matrix()


z1_arc = pairwise_col_diff(exp_GSE36059_arc) %>% as.matrix()
z3_arc = pairwise_col_diff(exp_GSE46474_arc) %>% as.matrix()
z2_arc = pairwise_col_diff(exp_GSE48581_arc) %>% as.matrix()



## log transform

exp_GSE36059_log <- exp_GSE36059
exp_GSE36059_log = exp_GSE36059_log + 1
exp_GSE36059_log = log(exp_GSE36059_log)

exp_GSE46474_log <- exp_GSE46474
exp_GSE46474_log = exp_GSE46474_log + 1
exp_GSE46474_log = log(exp_GSE46474_log)

exp_GSE48581_log <- exp_GSE48581
exp_GSE48581_log = exp_GSE48581_log + 1
exp_GSE48581_log = log(exp_GSE48581_log)

z1_log = exp_GSE36059_log  %>% as.matrix()
z3_log = exp_GSE46474_log %>% as.matrix()
z2_log = exp_GSE48581_log %>% as.matrix()

# View(z1_log)
outcome$characteristics_ch1.1
```

## pre transformation plot
```{r}
box11 = cbind(boxplot_tbl(z1, index = 1), GSE36059_id)
box22 = cbind(boxplot_tbl(z3, index = 1), GSE46474_id)
box33 = cbind(boxplot_tbl(z2, index = 1), GSE48581_id)
box4 = rbind(box11, box22, box33)

expressionplot <-
ggplot(data = box4, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  ylim(0,15) + 
  theme(legend.position="bottom") +
  theme(legend.title = element_blank()) +
  labs(title = "Raw Data") +
  theme(plot.title = element_text(size=10))

expressionplot
```

## Boxplot to visualise if the arc transformations were good
```{r}
box1_arc = cbind(boxplot_tbl(z1_arc, index = 1), GSE36059_id)
box2_arc = cbind(boxplot_tbl(z3_arc, index = 1), GSE46474_id)
box3_arc = cbind(boxplot_tbl(z2_arc, index = 1), GSE48581_id)
box4_arc = rbind(box1_arc, box2_arc, box3_arc)

arcplot <-
ggplot(data = box4_arc, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Arcsine transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

arcplot
```

## boxplot to see if the log transformation was good
```{r}
box1_log = cbind(boxplot_tbl(z1_log, index = 1), GSE36059_id)
box2_log = cbind(boxplot_tbl(z3_log, index = 1), GSE46474_id)
box3_log = cbind(boxplot_tbl(z2_log, index = 1), GSE48581_id)
box4_log = rbind(box1_log, box2_log, box3_log)

logplot <-
ggplot(data = box4_log, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Log transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

logplot
```
## getting the results vectors 
```{r}
p_GSE46474 = pData(GSE46474)
p_GSE48581 = pData(GSE48581) 
p_GSE36059 = pData(GSE36059) 


p_GSE36059$diagnosis = ifelse(p_GSE36059$characteristics_ch1 == "diagnosis: non-rejecting", 0, 1)
p_GSE48581$diagnosis = ifelse(p_GSE48581$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, 1)
p_GSE46474$diagnosis = ifelse(p_GSE46474$characteristics_ch1.5 == "procedure status: post-transplant non-rejection (NR)", 0, 1)

y1 = as.factor(p_GSE36059$diagnosis)
y2 = as.factor(p_GSE48581$diagnosis)
y3 = as.factor(p_GSE46474$diagnosis)
### GSE36059
### GSE48581
# these have reject + stable but categorized in more detail -> either have more groups that we are predicting, or we could do purely binary 
```


## running CPOP
```{r}
#CPOP model


counter = 0

bal_acc_y3_test = numeric(length(101))
features_test = numeric(length(101))
y3_alpha_accuracy_list = list()
features_list = list()

for (i in c(71)) {
  set.seed(3888)
  counter = counter + 1
  #alpha_vector = seq(0.01, 1, by=0.01)
  alpha_vector = c(1)
  counter2 = 0
  for(a in alpha_vector){
    counter2 = counter2 + 1
      cpopresult=
      cpop_model(
        z1_log,
        z2_log,
        y1,
        y2,
        w = NULL,
        n_features = 30,
        n_iter = 20,
        alpha = 1,
        family = "binomial",
        s = "lambda.min",
        cpop2_break = TRUE,
        cpop2_type = "sign",
        cpop2_mag = 1,
        cpop1_method = "normal",
        intercept = FALSE)

    prediction_z3 <- predict_cpop(cpopresult, z3_log, s = "lambda.min")
    
}
}
prediction_z3
y3
```


```{r}
## cross validation code
K = 5  # number of CV folds

# permute all the data, into 5 folds
n1 = nrow(z1_log)
n2 = nrow(z2_log)
cvSets_z1 = cvTools::cvFolds(n1, K) 
cvSets_z2 = cvTools::cvFolds(n2, K) 


cv_acc_z1 = NA  # initialise results vector
cv_acc_z2 = NA
cv_acc_alphas_z1 = NA
cv_acc_alphas_z2 = NA
for (j in 1:K) {
    test_id_z1 = cvSets_z1$subsets[cvSets_z1$which == j]
    test_id_z2 = cvSets_z2$subsets[cvSets_z2$which == j]
    z1_log_test = z1_log[test_id_z1, ]
    z1_log_train = z1_log[-test_id_z1, ]
    y1_test = y1[test_id_z1]
    y1_train = y1[-test_id_z1]
    z2_log_test = z2_log[test_id_z2, ]
    z2_log_train = z2_log[-test_id_z2, ]
    y2_test = y2[test_id_z2]
    y2_train = y2[-test_id_z2]
    

    counter = 0
    
    bal_acc_y3_test = numeric(length(101))
    features_test = numeric(length(101))
    y3_alpha_accuracy_list = list()
    features_list = list()
    
    for (i in c(71)) {
      set.seed(3888)
      counter = counter + 1
      alpha_vector = c(0.01, 0.5, 0.7, 0.9, 1)
      counter2 = 0
      for(a in alpha_vector){
        counter2 = counter2 + 1
          cpopresult=
          cpop_model(
            z1_log_train,
            z2_log_train,
            y1_train,
            y2_train,
            w = NULL,
            n_features = 30,
            n_iter = 20,
            alpha = 1,
            family = "binomial",
            s = "lambda.min",
            cpop2_break = TRUE,
            cpop2_type = "sign",
            cpop2_mag = 1,
            cpop1_method = "normal",
            intercept = FALSE)
    fit_z1 = predict_cpop(cpopresult, z1_log_test, s = "lambda.min")
    fit_z2 = predict_cpop(cpopresult, z2_log_test, s = "lambda.min")
    cv_acc_z1[j] = mean(fit_z1$cpop_model_avg_class == y1_test)
    cv_acc_z2[j] = mean(fit_z2$cpop_model_avg_class == y2_test)
    
    cv_acc_alphas_z1[counter2] = mean(cv_acc_z1)
    cv_acc_alphas_z2[counter2] = mean(cv_acc_z2)
      }
    }
}

cv_acc_z1
cv_acc_z2

cv_acc_alphas_z1
cv_acc_alphas_z2

## make this balanced acc
```


```{r}
results = cpopresult$coef_tbl
results$average = (results$coef1 +results$coef2)/2
results=results[2:nrow(results),]
results = results[order(-abs(results$average)),]
results = results[1:20,c(1,4)]

coef = results$coef_name
results$coef_name = sub(".*\\.", "", coef)

plot_lratio_network(results$coef_name, type = "visNetwork")

```


```{r}
#playing with network
results =
results %>%
  mutate(color = case_when(
    average > 0 ~ "blue",
    average < 0 ~ "red"
    ))
names = data.frame(feature1 = rep("",20),
                   feature2 = rep("",20),
                   coef_size = abs(results$average))
names
names$feature1 = sub("\\..*", "", coef) #getting the first node from the coef-name vector of the results df
names$feature2 = sub(".*--", "", coef) #getting the second node from the coef-name vector of the results df
names$color = results$color
network = plot_lratio_network(coef, type = "visNetwork")
network
names_uniq = unique(c(names$feature1, names$feature2))
names_uniq
numbers = names
numbers
for (a in 1:nrow(numbers)) {
  for (i in 1:length(names_uniq)) {
    if (numbers$feature2[a] == names_uniq[i]) {
      numbers$feature2[a] = i
    }
  }
}
for (a in 1:nrow(numbers)) {
  for (i in 1:length(names_uniq)) {
    if (numbers$feature1[a] == names_uniq[i]) {
      numbers$feature1[a] = i
    }
  }
}
numbers
clr = vector()
for (a in 1:length(names_uniq)) {
  for (i in 1:nrow(names)) {
    if (names_uniq[a] == names[i,1] | names_uniq[a] == names[i,2]) {
      clr[a] = names$color[i]
    }
  }
}
clr
names
edges = data.frame(from = numbers$feature1, to = numbers$feature2, value = names$coef_size)
nodes = data.frame(id = c(1:length(names_uniq)), 
                   label = names_uniq, 
                   color = clr)
visNetwork(nodes, edges, height = "500px", width = "100%")

```
```

```{r}
## Random CPOP code while I try debug earlier chunk
    c = is.na(prediction_z3$cpop_model_avg_class)
    na_index1 = which(c %in% c(TRUE))
    
    if (length(na_index1) == 0) {
      bal_acc <- yardstick::bal_accuracy_vec(
      truth = y3,
      estimate = factor(coalesce(prediction_z3$cpop_model_avg_class, "0")))
    }
    else {
      bal_acc <- yardstick::bal_accuracy_vec(
      truth = y3,
      estimate = factor(coalesce(prediction_z3$cpop_model_avg_class, "0")))
    }
    
    bal_acc_y3_test[counter2] = bal_acc
    features_test[counter2] = length(cpopresult$features)
  }
  
  y3_alpha_accuracy_list[[counter]] = bal_acc_y3_test
  features_list[[counter]] = features_test
  print(y3_alpha_accuracy_list)
```


```{r}
## Khang play area

data(cpop_data_binary, package = 'CPOP')
x1 = cpop_data_binary$x1
x2 = cpop_data_binary$x2
x3 = cpop_data_binary$x3
y1 = cpop_data_binary$y1
y2 = cpop_data_binary$y2
y3 = cpop_data_binary$y3




```

