---
title: "KNN"
output: html_notebook
---

This notebook contains research on using KNN for the Kidney Graft Genetics Project to predict graft survival.

```{r}
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(class)
library(cvTools)
library(randomForest)
library(GEOquery) 
library(R.utils)
library(reshape2)
library(limma)
library(dplyr)
library(e1071)
library(DT)
library(viridis)
library(plotly)
library(scales)
library(CPOP) #devtools::install_github("sydneybiox/CPOP") #Will take a while to download and requires user input in console
library(matrixStats)
```

```{r}
gse36059 = getGEO("GSE36059")[[1]]
gse48581 = getGEO("GSE48581")[[1]]
gse129166 = getGEO("GSE129166")[[1]]
```



```{r}
gse36059_f = fData(gse36059)
gse36059_f
gse48581_f = fData(gse48581)
gse48581_f
gse129166_f = fData(gse129166)
gse129166_f
```

```{r}
gse36059_p = pData(gse36059)
gse36059_p
gse48581_p = pData(gse48581)
gse48581_p
gse129166_p = pData(gse129166)
gse129166_p
```
```{r}
#Remove Nephrectomy outcomes
gse36059_p = gse36059_p[!(gse36059_p$characteristics_ch1=="diagnosis: Nephrectomy"),]

gse48581_p = gse48581_p[!(gse48581_p$characteristics_ch1=="diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): nephrectomy"),]
```


```{r}
#Encodes stable as 0 and rejecting as 1
gse36059_p$diagnosis = ifelse(gse36059_p$characteristics_ch1 == "diagnosis: non-rejecting", 0, 1)
gse48581_p$diagnosis = ifelse(gse48581_p$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, 1)
gse129166_p$diagnosis = ifelse((gse129166_p$characteristics_ch1.1 == "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0"), 0, 1)
```


```{r}
#Encodes stable as 0, ABMR as 1, TCMR as 2, Mixed as 3
gse36059_p$exact_diagnosis = ifelse(gse36059_p$characteristics_ch1 == "diagnosis: non-rejecting", 0, ifelse(gse36059_p$characteristics_ch1 =="diagnosis: ABMR", 1, ifelse(gse36059_p$characteristics_ch1 == "diagnosis: TCMR", 2, 3)))

gse48581_p$exact_diagnosis = ifelse(gse48581_p$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, ifelse(gse48581_p$characteristics_ch1.1 =="diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): ABMR", 1, ifelse(gse48581_p$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): TCMR", 2, 3)))

gse129166_p$exact_diagnosis = ifelse((gse129166_p$characteristics_ch1.1 == "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0"), 0, ifelse((gse129166_p$characteristics_ch1.1 != "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 != "abmr (no: 0_Yes:1): 0"), 3, ifelse((gse129166_p$characteristics_ch1.1 != "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0"), 2, 1)))
```


```{r}
gse36059_p
gse48581_p
gse129166_p
```
Main outcomes are non-rejecting, TCMR (acute T-cell–mediated rejection), ABMR (anti-donor antibody-mediated rejection), and MIXED. Also has nephrectomy.


```{r}
gse36059_ex = data.frame(t(exprs(gse36059)))
gse36059_ex
gse48581_ex = data.frame(t(exprs(gse48581)))
gse48581_ex
gse129166_ex = data.frame(t(exprs(gse129166)))
gse129166_ex
```
```{r}
start = 1
stop = 100
boxplot(gse36059_ex[start:stop])
boxplot(gse48581_ex[start:stop])
boxplot(gse129166_ex[start:stop])
```

```{r}
#TODO 
pairwise_preprocess = function(GSE, top_x) {
  exp_GSE = (exprs(GSE))
  Variance = rowVars(as.matrix(exp_GSE))
  Variance = as.data.frame(Variance)
  exp_GSE = as.data.frame(exp_GSE)
  exp_GSE = cbind(exp_GSE, variance = Variance)
  exp_GSE = slice_max(exp_GSE, order_by = Variance, n = top_x)
  exp_GSE = subset(exp_GSE, select = -c(Variance))
  row_names_exp_GSE = rownames(exp_GSE)
  
  return(exp_GSE)
}

pairwise = function(exp_GSE, intersection, transform_type) {
  
  exp_GSE = as.data.frame(t(as.matrix(exp_GSE)))
  exp_GSE = subset(exp_GSE, select = c(intersection))
  
  z = exp_GSE
  
  if (transform_type == "Arc") {
    z = z / max(z)
    z = asin(sqrt(z))
    z = pairwise_col_diff(z) %>% as.matrix()
  }
  else if (transform_type == "Log") {
    z = z + 1
    z = log(z)
  }
  else if (transform_type == "Pair"){
    z = z %>% as.matrix()
    z_pairwise = pairwise_col_diff(z) %>% as.matrix()
    
  }
  
  return(z)
  
}

z1 = pairwise_preprocess(gse36059, 2000)
z2 = pairwise_preprocess(gse48581, 2000)

intersection = intersect(rownames(z1), rownames(z2))

z1_pair = pairwise(z1, intersection, "Log")
z2_pair = pairwise(z2, intersection, "Log")

```
```{r}
boxplot(z1_pair[1:50,], ylim = c(0, 4))
```



```{r}
# TODO: Export processed code for all datasets

```

```{r}
# TODO: Find overlap in features selected in KNN for the different datasets
```


```{r}
#CPOP data

## keeping only the 100 most variable genes in my data frame 
exp_GSE36059 = (exprs(gse36059))
Variance = rowVars(as.matrix(exp_GSE36059))
Variance = as.data.frame(Variance)
exp_GSE36059 = as.data.frame(exp_GSE36059)
exp_GSE36059 = cbind(exp_GSE36059, variance = Variance)
exp_GSE36059 = slice_max(exp_GSE36059, order_by = Variance, n = 2000)
exp_GSE36059 = subset(exp_GSE36059, select = -c(Variance))
row_names_exp_GSE36074 = rownames(exp_GSE36059)

exp_GSE48581 = (exprs(gse48581))
Variance = rowVars(as.matrix(exp_GSE48581))
Variance = as.data.frame(Variance)
exp_GSE48581 = as.data.frame(exp_GSE48581)
exp_GSE48581 = cbind(exp_GSE48581, variance = Variance)
exp_GSE48581 = slice_max(exp_GSE48581, order_by = Variance, n = 2000)
exp_GSE48581 = subset(exp_GSE48581, select = -c(Variance))
row_names_exp_GSE48581 = rownames(exp_GSE48581)

exp_GSE129166 = (exprs(gse129166))
Variance = rowVars(as.matrix(exp_GSE129166))
Variance = as.data.frame(Variance)
exp_GSE129166 = as.data.frame(exp_GSE129166)
exp_GSE129166 = cbind(exp_GSE129166, variance = Variance)
exp_GSE129166 = slice_max(exp_GSE129166, order_by = Variance, n = 2000)
exp_GSE129166 = subset(exp_GSE129166, select = -c(Variance))
row_names_exp_GSE129166 = rownames(exp_GSE129166)
```

```{r}
# Takes overlap
intersection = intersect(row_names_exp_GSE34748, row_names_exp_GSE46474)
intersection = intersect(intersection, row_names_exp_GSE129166)

exp_GSE34748 = as.data.frame(t(as.matrix(exp_GSE34748)))
exp_GSE34748 = subset(exp_GSE34748, select = c(intersection))

exp_GSE46474 = as.data.frame(t(as.matrix(exp_GSE46474)))
exp_GSE46474 = subset(exp_GSE46474, select = c(intersection))

exp_GSE129166 = as.data.frame(t(as.matrix(exp_GSE129166)))
exp_GSE129166 = subset(exp_GSE129166, select = c(intersection))

GSE34748_id <- data.frame("Dataset" = rep("GSE34748",nrow(exp_GSE34748)))
GSE46474_id <- data.frame("Dataset" = rep("GSE46474",nrow(exp_GSE46474)))
GSE129166_id <- data.frame("Dataset" = rep("GSE129166",nrow(exp_GSE129166)))

z1 = exp_GSE34748 %>% as.matrix()
z2 = exp_GSE46474 %>% as.matrix()
z3 = exp_GSE129166 %>% as.matrix()

## arcsine transformation

exp_GSE34748_arc <- exp_GSE34748
exp_GSE34748_arc = exp_GSE34748_arc / max(exp_GSE34748_arc)
exp_GSE34748_arc = asin(sqrt(exp_GSE34748_arc))

exp_GSE46474_arc <- exp_GSE46474
exp_GSE46474_arc = exp_GSE46474_arc / max(exp_GSE46474_arc)
exp_GSE46474_arc = asin(sqrt(exp_GSE46474_arc))

exp_GSE129166_arc <- exp_GSE129166
exp_GSE129166_arc = exp_GSE129166_arc / max(exp_GSE129166_arc)
exp_GSE129166_arc = asin(sqrt(exp_GSE129166_arc))

z1_pairwise = pairwise_col_diff(z1) %>% as.matrix()
z2_pairwise = pairwise_col_diff(z2) %>% as.matrix()
z3_pairwise = pairwise_col_diff(z3) %>% as.matrix()


z1_arc = pairwise_col_diff(exp_GSE34748_arc) %>% as.matrix()
z2_arc = pairwise_col_diff(exp_GSE46474_arc) %>% as.matrix()
z3_arc = pairwise_col_diff(exp_GSE129166_arc) %>% as.matrix()

## log transform

exp_GSE34748_log <- exp_GSE34748
exp_GSE34748_log = exp_GSE34748_log + 1
exp_GSE34748_log = log(exp_GSE34748_log)

exp_GSE46474_log <- exp_GSE46474
exp_GSE46474_log = exp_GSE46474_log + 1
exp_GSE46474_log = log(exp_GSE46474_log)

exp_GSE129166_log <- exp_GSE129166
exp_GSE129166_log = exp_GSE129166_log + 1
exp_GSE129166_log = log(exp_GSE129166_log)

z1_log = pairwise_col_diff(exp_GSE34748_log) %>% as.matrix()
z2_log = pairwise_col_diff(exp_GSE46474_log) %>% as.matrix()
z3_log = pairwise_col_diff(exp_GSE129166_log) %>% as.matrix()


```

## pre transformation plot
```{r}
box11 = cbind(boxplot_tbl(z1, index = 1), GSE34748_id)
box22 = cbind(boxplot_tbl(z2, index = 1), GSE46474_id)
box33 = cbind(boxplot_tbl(z3, index = 1), GSE129166_id)
box4 = rbind(box11, box22, box33)

expressionplot <-
ggplot(data = box4, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  ylim(0,15) + 
  theme(legend.position="bottom") +
  theme(legend.title = element_blank()) +
  labs(title = "Raw Data") +
  theme(plot.title = element_text(size=10))

expressionplot
```

## Boxplot to visualise if the arc transformations were good
```{r}
box1_arc = cbind(boxplot_tbl(z1_arc, index = 1), GSE34748_id)
box2_arc = cbind(boxplot_tbl(z2_arc, index = 1), GSE46474_id)
box3_arc = cbind(boxplot_tbl(z3_arc, index = 1), GSE129166_id)
box4_arc = rbind(box1_arc, box2_arc, box3_arc)

arcplot <-
ggplot(data = box4_arc, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Arcsine transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

arcplot
```

## boxplot to see if the log transformation was good
```{r}
box1_log = cbind(boxplot_tbl(z1_log, index = 1), GSE34748_id)
box2_log = cbind(boxplot_tbl(z2_log, index = 1), GSE46474_id)
box3_log = cbind(boxplot_tbl(z3_log, index = 1), GSE129166_id)
box4_log = rbind(box1_log, box2_log, box3_log)

logplot <-
ggplot(data = box4_log, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Log transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

logplot
```
## getting the results vectors 
```{r}
pData(GSE46474)
pData(GSE129166) #is there rejection and stable
pData(GSE34748) #no reject or stable



### GSE36059
### GSE48581
# these have reject + stable but categorized in more detail -> either have more groups that we are predicting, or we could do purely binary 
```