---
title: "Liam Research 1"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(class)
library(cvTools)
library(randomForest)
library(GEOquery) 
library(R.utils)
library(reshape2)
library(limma)
library(dplyr)
library(e1071)
library(DT)
library(viridis)
library(plotly)
library(scales)
```

```{r}
geo_name_list = c("GSE129166", "GSE34748", "GSE51675", "GSE15296", "GSE46474", "GSE50084")
geo_list = c(getGEO(filename="Data/GSE129166_series_matrix.txt.gz"), getGEO(filename="Data/GSE34748_series_matrix.txt.gz"), getGEO(filename="Data/GSE51675_series_matrix.txt.gz"), getGEO(filename="Data/GSE15296_series_matrix.txt.gz"), getGEO(filename="Data/GSE46474_series_matrix.txt.gz"), getGEO(filename="Data/GSE50084_series_matrix.txt.gz"))
```

```{r}
for(geo in geo_list) {
  print(fData(geo))
  print(names(which(colSums(is.na(fData(geo)))>0)))
}
```
All gse have a gene symbol column which can be used to match records except for the last one, GSE50084 which has the gene as the 2nd listed element under gene_assignment, and GSE51675 which is missing all of its gene info. 


```{r}
for(geo in geo_list) {
  print(pData(geo))
  print(names(which(colSums(is.na(pData(geo)))>0)))
}
```

```{r}
for (geo in geo_list) {
  print(data.frame(t(exprs(geo))))
  print(names(which(colSums(is.na(data.frame(t(exprs(geo)))))>0)))
}
```

```{r}
# TODO: Find the outlier genes and compare
for (geo in geo_list) {
  df = data.frame(t(exprs(geo)))
  boxplot(df[1:100])
}
```
Plots for GSE51675 and GSE15296 show that they have been transformed somehow. Other than those, the other 4 seem fairly contained between 0-14 expression intensity.

GSE51675 has been discarded due to small sample size, missing info, and transformed expression values. GSE50084 has been discarded



```{r}
# Contains "GSE129166", "GSE34748"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[2]])["Gene Symbol"])[[1]])
```

```{r}
# Contains "GSE129166", "GSE46474"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[5]])["Gene Symbol"])[[1]])
```

All 3 have perfect overlap so the datasets recommended for use are "GSE129166", "GSE34748", and "GSE46474"

```{r}
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[4]])["Gene Symbol"])[[1]])
```
