---
title: "Liam Research 1"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(class)
library(cvTools)
library(randomForest)
library(GEOquery) 
library(R.utils)
library(reshape2)
library(limma)
library(dplyr)
library(e1071)
library(DT)
library(viridis)
library(plotly)
library(scales)
library(CPOP)
library(matrixStats)
library(visNetwork)
```

```{r}
geo_name_list = c("GSE129166", "GSE34748", "GSE51675", "GSE15296", "GSE46474", "GSE50084", "GSE36059", "GSE48581")
geo_list = c(getGEO(filename="Data/GSE129166_series_matrix.txt.gz"), getGEO(filename="Data/GSE34748_series_matrix.txt.gz"), getGEO(filename="Data/GSE51675_series_matrix.txt.gz"), getGEO(filename="Data/GSE15296_series_matrix.txt.gz"), getGEO(filename="Data/GSE46474_series_matrix.txt.gz"), getGEO(filename="Data/GSE50084_series_matrix.txt.gz"), getGEO("GSE36059"), getGEO("GSE48581"))

### GSE36059
### GSE48581

getGEO(filename="Data/GSE36059_series_matrix.txt.gz")
```

```{r}
for(geo in geo_list) {
  print(fData(geo))
  print(names(which(colSums(is.na(fData(geo)))>0)))
}
```
All gse have a gene symbol column which can be used to match records except for the last one, GSE50084 which has the gene as the 2nd listed element under gene_assignment, and GSE51675 which is missing all of its gene info. 


```{r}
for(geo in geo_list) {
  print(pData(geo))
  print(names(which(colSums(is.na(pData(geo)))>0)))
}
```

```{r}
for (geo in geo_list) {
  print(data.frame(t(exprs(geo))))
  print(names(which(colSums(is.na(data.frame(t(exprs(geo)))))>0)))
}
```

```{r}
for (geo in geo_list) {
  df = data.frame(t(exprs(geo)))
  boxplot(df[1:100])
}
```
Plots for GSE51675 and GSE15296 show that they have been transformed somehow. Other than those, the other 4 seem fairly contained between 0-14 expression intensity.

GSE51675 has been discarded due to small sample size, missing info, and transformed expression values. GSE50084 has been discarded



```{r}
# Contains "GSE129166", "GSE34748"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[2]])["Gene Symbol"])[[1]])
```

```{r}
# Contains "GSE129166", "GSE46474"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[5]])["Gene Symbol"])[[1]])
```

All 3 have perfect overlap so the datasets recommended for use are "GSE129166", "GSE34748", and "GSE46474"

(Mukund): Also just thinking, but we probably do want to go through the effort of trying to clean another dataset to match these. firstly for the sake of increased difficulty in the eyes of Jean, but also sometimes without going through the process it can be easy to overlook something in the data cleaning aspect. I am just a bit worried because after talking to some other people they have all told me the data cleaning for this is really hard so not sure if we are actually missing something. 

```{r}
GSE34748 = getGEO(filename="Data/GSE34748_series_matrix.txt.gz")
clinical_GSE34748 = pData(GSE34748)
clinical_GSE34748
```
Also just doing some checking and some of these don't actually have clinical data

```{r}
GSE46474 = getGEO(filename="Data/GSE46474_series_matrix.txt.gz")
clinical_GSE46474 = pData(GSE46474)
clinical_GSE46474
```
This one has clinical data!!
```{r}
GSE129166 = getGEO(filename="Data/GSE129166_series_matrix.txt.gz")
clinical_GSE129166 = pData(GSE129166)
clinical_GSE129166
```
Also no clinical data
```{r}
#GSE129166 = getGEO(filename="Data/GSE129166_series_matrix.txt.gz")
#clinical_GSE129166 = pData(GSE129166)

```

## Mukund CPOP analysis 
```{r}
#CPOP data

GSE36059 = getGEO("GSE36059")[[1]]
GSE48581 = getGEO("GSE48581")[[1]]
GSE129166 = getGEO("GSE129166")[[1]]
#GSE15296 = getGEO("GSE15296")[[1]]


```


```{r}
## keeping only the 100 most variable genes in my data frame 
exp_GSE36059 = (exprs(GSE36059)) ##Biopsy
Variance = rowVars(as.matrix(exp_GSE36059))
Variance = as.data.frame(Variance)
exp_GSE36059 = as.data.frame(exp_GSE36059)
exp_GSE36059 = cbind(exp_GSE36059, variance = Variance)
exp_GSE36059 = slice_max(exp_GSE36059, order_by = Variance, n = 2000)
exp_GSE36059 = subset(exp_GSE36059, select = -c(Variance))
row_names_exp_GSE36059 = rownames(exp_GSE36059)



exp_GSE46474 = (exprs(GSE46474)) #blood
Variance = rowVars(as.matrix(exp_GSE46474))
Variance = as.data.frame(Variance)
exp_GSE46474 = as.data.frame(exp_GSE46474)
exp_GSE46474 = cbind(exp_GSE46474, variance = Variance)
exp_GSE46474 = slice_max(exp_GSE46474, order_by = Variance, n = 2000)
exp_GSE46474 = subset(exp_GSE46474, select = -c(Variance))
row_names_exp_GSE46474 = rownames(exp_GSE46474)


exp_GSE48581 = (exprs(GSE48581)) #biopsy
Variance = rowVars(as.matrix(exp_GSE48581))
Variance = as.data.frame(Variance)
exp_GSE48581 = as.data.frame(exp_GSE48581)
exp_GSE48581 = cbind(exp_GSE48581, variance = Variance)
exp_GSE48581 = slice_max(exp_GSE48581, order_by = Variance, n = 2000)
exp_GSE48581 = subset(exp_GSE48581, select = -c(Variance))
row_names_exp_GSE48581 = rownames(exp_GSE48581)

#129166
exp_GSE129166 = (exprs(GSE129166)) #blood
Variance = rowVars(as.matrix(exp_GSE129166))
Variance = as.data.frame(Variance)
exp_GSE129166 = as.data.frame(exp_GSE129166)
exp_GSE129166 = cbind(exp_GSE129166, variance = Variance)
exp_GSE129166 = slice_max(exp_GSE129166, order_by = Variance, n = 2000)
exp_GSE129166 = subset(exp_GSE129166, select = -c(Variance))
row_names_exp_GSE129166 = rownames(exp_GSE129166)

intersection = intersect(row_names_exp_GSE36059, row_names_exp_GSE46474)
intersection = intersect(intersection, row_names_exp_GSE48581)
intersection = intersect(intersection, row_names_exp_GSE129166)

exp_GSE36059 = as.data.frame(t(as.matrix(exp_GSE36059)))
exp_GSE36059 = subset(exp_GSE36059, select = c(intersection))

exp_GSE46474 = as.data.frame(t(as.matrix(exp_GSE46474)))
exp_GSE46474 = subset(exp_GSE46474, select = c(intersection))

exp_GSE48581 = as.data.frame(t(as.matrix(exp_GSE48581)))
exp_GSE48581 = subset(exp_GSE48581, select = c(intersection))

exp_GSE129166 = as.data.frame(t(as.matrix(exp_GSE129166)))
exp_GSE129166 = subset(exp_GSE129166, select = c(intersection))

GSE36059_id <- data.frame("Dataset" = rep("GSE36059",nrow(exp_GSE36059)))
GSE46474_id <- data.frame("Dataset" = rep("GSE46474",nrow(exp_GSE46474)))
GSE48581_id <- data.frame("Dataset" = rep("GSE48581",nrow(exp_GSE48581)))
GSE129166_id <- data.frame("Dataset" = rep("GSE129166",nrow(exp_GSE129166)))

z1 = exp_GSE36059 %>% as.matrix()
z3 = exp_GSE46474 %>% as.matrix()
z2 = exp_GSE48581 %>% as.matrix()
z4 = exp_GSE129166 %>% as.matrix()
## arcsine transformation

exp_GSE36059_arc <- exp_GSE36059
exp_GSE36059_arc = exp_GSE36059_arc / max(exp_GSE36059_arc)
exp_GSE36059_arc = asin(sqrt(exp_GSE36059_arc))

exp_GSE46474_arc <- exp_GSE46474
exp_GSE46474_arc = exp_GSE46474_arc / max(exp_GSE46474_arc)
exp_GSE46474_arc = asin(sqrt(exp_GSE46474_arc))

exp_GSE48581_arc <- exp_GSE48581
exp_GSE48581_arc = exp_GSE48581_arc / max(exp_GSE48581_arc)
exp_GSE48581_arc = asin(sqrt(exp_GSE48581_arc))

z1_pairwise = pairwise_col_diff(z1) %>% as.matrix()
z2_pairwise = pairwise_col_diff(z2) %>% as.matrix()
z3_pairwise = pairwise_col_diff(z3) %>% as.matrix()


z1_arc = pairwise_col_diff(exp_GSE36059_arc) %>% as.matrix()
z3_arc = pairwise_col_diff(exp_GSE46474_arc) %>% as.matrix()
z2_arc = pairwise_col_diff(exp_GSE48581_arc) %>% as.matrix()

## log transform

exp_GSE36059_log <- exp_GSE36059
exp_GSE36059_log = exp_GSE36059_log + 1
exp_GSE36059_log = log(exp_GSE36059_log)

exp_GSE46474_log <- exp_GSE46474
exp_GSE46474_log = exp_GSE46474_log + 1
exp_GSE46474_log = log(exp_GSE46474_log)

exp_GSE48581_log <- exp_GSE48581
exp_GSE48581_log = exp_GSE48581_log + 1
exp_GSE48581_log = log(exp_GSE48581_log)

exp_GSE129166_log <- exp_GSE129166
exp_GSE129166_log = exp_GSE129166_log + 1
exp_GSE129166_log = log(exp_GSE129166_log)

z1_log = pairwise_col_diff(exp_GSE36059_log) %>% as.matrix()
z3_log = pairwise_col_diff(exp_GSE46474_log) %>% as.matrix()
z2_log = pairwise_col_diff(exp_GSE48581_log) %>% as.matrix()
z4_log = pairwise_col_diff(exp_GSE129166_log) %>% as.matrix()


```

## pre transformation plot
```{r}
box11 = cbind(boxplot_tbl(z1, index = 1), GSE36059_id)
box22 = cbind(boxplot_tbl(z3, index = 1), GSE46474_id)
box33 = cbind(boxplot_tbl(z2, index = 1), GSE48581_id)
box4 = rbind(box11, box22, box33)

expressionplot <-
ggplot(data = box4, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  ylim(0,15) + 
  theme(legend.position="bottom") +
  theme(legend.title = element_blank()) +
  labs(title = "Raw Data") +
  theme(plot.title = element_text(size=10))

expressionplot
```

## Boxplot to visualise if the arc transformations were good
```{r}
box1_arc = cbind(boxplot_tbl(z1_arc, index = 1), GSE36059_id)
box2_arc = cbind(boxplot_tbl(z3_arc, index = 1), GSE46474_id)
box3_arc = cbind(boxplot_tbl(z2_arc, index = 1), GSE48581_id)
box4_arc = rbind(box1_arc, box2_arc, box3_arc)

arcplot <-
ggplot(data = box4_arc, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Arcsine transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

arcplot
```

## boxplot to see if the log transformation was good
```{r}
box1_log = cbind(boxplot_tbl(z1_log, index = 1), GSE36059_id)
box2_log = cbind(boxplot_tbl(z4_log, index = 1), GSE129166_id)
box3_log = cbind(boxplot_tbl(z2_log, index = 1), GSE48581_id)
box4_log = rbind(box1_log, box2_log, box3_log)

logplot <-
ggplot(data = box4_log, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Log transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

logplot
```
## getting the results vectors for only reject or stable
```{r}
p_GSE46474 = pData(GSE46474)
p_GSE48581 = pData(GSE48581) 
p_GSE36059 = pData(GSE36059) 

p_GSE36059$diagnosis = ifelse(p_GSE36059$characteristics_ch1 == "diagnosis: non-rejecting", 0, 1)
p_GSE48581$diagnosis = ifelse(p_GSE48581$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, 1)
p_GSE46474$diagnosis = ifelse(p_GSE46474$characteristics_ch1.5 == "procedure status: post-transplant non-rejection (NR)", 0, 1)

y1 = as.factor(p_GSE36059$diagnosis)
y2 = as.factor(p_GSE48581$diagnosis)
y3 = as.factor(p_GSE46474$diagnosis)
### GSE36059
### GSE48581
# these have reject + stable but categorized in more detail -> either have more groups that we are predicting, or we could do purely binary 
```

## getting results vector for more 
```{r}
gse36059_p$exact_diagnosis = ifelse(gse36059_p$characteristics_ch1 == "diagnosis: non-rejecting", 0, ifelse(gse36059_p$characteristics_ch1 =="diagnosis: ABMR", 1, ifelse(gse36059_p$characteristics_ch1 == "diagnosis: TCMR", 2, 3)))

gse48581_p$exact_diagnosis = ifelse(gse48581_p$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, ifelse(gse48581_p$characteristics_ch1.1 =="diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): ABMR", 1, ifelse(gse48581_p$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): TCMR", 2, 3)))

gse129166_p$exact_diagnosis = ifelse((gse129166_p$characteristics_ch1.1 == "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0"), 0, ifelse((gse129166_p$characteristics_ch1.1 != "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 != "abmr (no: 0_Yes:1): 0"), 3, ifelse((gse129166_p$characteristics_ch1.1 != "tcmr (no: 0_borderline:1_TCMR:2): 0") & (gse129166_p$characteristics_ch1.2 == "abmr (no: 0_Yes:1): 0"), 2, 1)))
```


## running CPOP
```{r}
#CPOP model


counter = 0

bal_acc_y3_test = numeric(length(101))
features_test = numeric(length(101))
y3_alpha_accuracy_list = list()
features_list = list()

for (i in c(71)) {
  set.seed(3888)
  counter = counter + 1
  #alpha_vector = seq(0.01, 1, by=0.01)
  alpha_vector = c(1)
  counter2 = 0
  for(a in alpha_vector){
    counter2 = counter2 + 1
      cpopresult=
      cpop_model(
        z1_log,
        z2_log,
        y1,
        y2,
        w = NULL,
        n_features = 30,
        n_iter = 20,
        alpha = 1,
        family = "binomial",
        s = "lambda.min",
        cpop2_break = TRUE,
        cpop2_type = "sign",
        cpop2_mag = 1,
        cpop1_method = "normal",
        intercept = FALSE)

    prediction_z3 <- predict_cpop(cpopresult, z3_log, s = "lambda.min")
    
}
}
prediction_z3
y3
```


```{r}
## cross validation code
K = 5  # number of CV folds

# permute all the data, into 5 folds
n1 = nrow(z1_log)
n2 = nrow(z2_log)
cvSets_z1 = cvTools::cvFolds(n1, K) 
cvSets_z2 = cvTools::cvFolds(n2, K) 


cv_acc_z1 = NA  # initialise results vector
cv_acc_z2 = NA
cv_acc_alphas_z1 = NA
cv_acc_alphas_z2 = NA
for (j in 1:K) {
    test_id_z1 = cvSets_z1$subsets[cvSets_z1$which == j]
    test_id_z2 = cvSets_z2$subsets[cvSets_z2$which == j]
    z1_log_test = z1_log[test_id_z1, ]
    z1_log_train = z1_log[-test_id_z1, ]
    y1_test = y1[test_id_z1]
    y1_train = y1[-test_id_z1]
    z2_log_test = z2_log[test_id_z2, ]
    z2_log_train = z2_log[-test_id_z2, ]
    y2_test = y2[test_id_z2]
    y2_train = y2[-test_id_z2]
    

    counter = 0
    
    bal_acc_y3_test = numeric(length(101))
    features_test = numeric(length(101))
    y3_alpha_accuracy_list = list()
    features_list = list()
    
    for (i in c(71)) {
      set.seed(3888)
      counter = counter + 1
      #alpha_vector = seq(0.01, 1, by=0.01)
      alpha_vector = c(0.01, 0.5, 0.7, 0.9, 1)
      counter2 = 0
      for(a in alpha_vector){
        counter2 = counter2 + 1
          cpopresult=
          cpop_model(
            z1_log_train,
            z2_log_train,
            y1_train,
            y2_train,
            w = NULL,
            n_features = 30,
            n_iter = 20,
            alpha = 1,
            family = "binomial",
            s = "lambda.min",
            cpop2_break = TRUE,
            cpop2_type = "sign",
            cpop2_mag = 1,
            cpop1_method = "normal",
            intercept = FALSE)
    fit_z1 = predict_cpop(cpopresult, z1_log_test, s = "lambda.min")
    fit_z2 = predict_cpop(cpopresult, z2_log_test, s = "lambda.min")
    cv_acc_z1[j] = mean(fit_z1$cpop_model_avg_class == y1_test)
    cv_acc_z2[j] = mean(fit_z2$cpop_model_avg_class == y2_test)
    
    cv_acc_alphas_z1[counter2] = mean(cv_acc_z1)
    cv_acc_alphas_z2[counter2] = mean(cv_acc_z2)
      }
    }
}

cv_acc_z1
cv_acc_z2

cv_acc_alphas_z1
cv_acc_alphas_z2

## make this balanced acc
```

#pre-lim work to make feature graph 
```{r}
results = cpopresult$coef_tbl
results$average = (results$coef1 +results$coef2)/2
results=results[2:nrow(results),]
results = results[order(-abs(results$average)),]
results = results[1:20,c(1,4)]

coef = results$coef_name
results$coef_name = sub(".*\\.", "", coef)

plot_lratio_network(results$coef_name, type = "visNetwork")
results
```

#making the final feature graph with colours and sizing. Make colours better + add explanation about what the colour + size means. Change the size of the edges! 
```{r}
#playing with network
results =
results %>%
  mutate(color = case_when(
    average > 0 ~ "blue",
    average < 0 ~ "red"
    ))

names = data.frame(feature1 = rep("",20),
                   feature2 = rep("",20),
                   coef_size = abs(results$average))
names

names$feature1 = sub("\\..*", "", coef) #getting the first node from the coef-name vector of the results df
names$feature2 = sub(".*--", "", coef) #getting the second node from the coef-name vector of the results df
names$color = results$color


network = plot_lratio_network(coef, type = "visNetwork")
network
names_uniq = unique(c(names$feature1, names$feature2))
names_uniq
numbers = names
numbers
for (a in 1:nrow(numbers)) {
  for (i in 1:length(names_uniq)) {
    if (numbers$feature2[a] == names_uniq[i]) {
      numbers$feature2[a] = i
    }
  }
}

for (a in 1:nrow(numbers)) {
  for (i in 1:length(names_uniq)) {
    if (numbers$feature1[a] == names_uniq[i]) {
      numbers$feature1[a] = i
    }
  }
}
numbers
clr = vector()
for (a in 1:length(names_uniq)) {
  for (i in 1:nrow(names)) {
    if (names_uniq[a] == names[i,1] | names_uniq[a] == names[i,2]) {
      clr[a] = names$color[i]
    }
  }
}
clr
names
edges = data.frame(from = numbers$feature1, to = numbers$feature2, value = names$coef_size)
nodes = data.frame(id = c(1:35), 
                   label = names_uniq, 
                   color = clr)

visNetwork(nodes, edges, height = "500px", width = "100%")


length(names_uniq)

```


Dirty data fixing 
```{r}

## any value that isnt 
dirty = read.csv("Data/dirtyuser.csv")
head(dirty)
dummy_col = dirty$GSM3701704
for (i in c(1:length(dummy_col))){
  if (is.na(as.double(dummy_col[i])) == TRUE){
    dummy_col[i] = NA
  }
  else if (dummy_col[i] < 0){
    dummy_col[i] = NA
  }
}

```

