---
title: "Liam Research 1"
output: html_notebook
---

```{r, include=FALSE}
library(tidyverse)
library(tuneR)
library(devtools)
library(ggplot2)
library(tsfeatures)
library(class)
library(cvTools)
library(randomForest)
library(GEOquery) 
library(R.utils)
library(reshape2)
library(limma)
library(dplyr)
library(e1071)
library(DT)
library(viridis)
library(plotly)
library(scales)
library(CPOP)
library(matrixStats)
```

```{r}
geo_name_list = c("GSE129166", "GSE34748", "GSE51675", "GSE15296", "GSE46474", "GSE50084", "GSE36059", "GSE48581")
geo_list = c(getGEO(filename="Data/GSE129166_series_matrix.txt.gz"), getGEO(filename="Data/GSE34748_series_matrix.txt.gz"), getGEO(filename="Data/GSE51675_series_matrix.txt.gz"), getGEO(filename="Data/GSE15296_series_matrix.txt.gz"), getGEO(filename="Data/GSE46474_series_matrix.txt.gz"), getGEO(filename="Data/GSE50084_series_matrix.txt.gz"), getGEO("GSE36059"), getGEO("GSE48581"))

### GSE36059
### GSE48581

getGEO(filename="Data/GSE36059_series_matrix.txt.gz")
```

```{r}
for(geo in geo_list) {
  print(fData(geo))
  print(names(which(colSums(is.na(fData(geo)))>0)))
}
```
All gse have a gene symbol column which can be used to match records except for the last one, GSE50084 which has the gene as the 2nd listed element under gene_assignment, and GSE51675 which is missing all of its gene info. 


```{r}
for(geo in geo_list) {
  print(pData(geo))
  print(names(which(colSums(is.na(pData(geo)))>0)))
}
```

```{r}
for (geo in geo_list) {
  print(data.frame(t(exprs(geo))))
  print(names(which(colSums(is.na(data.frame(t(exprs(geo)))))>0)))
}
```

```{r}
for (geo in geo_list) {
  df = data.frame(t(exprs(geo)))
  boxplot(df[1:100])
}
```
Plots for GSE51675 and GSE15296 show that they have been transformed somehow. Other than those, the other 4 seem fairly contained between 0-14 expression intensity.

GSE51675 has been discarded due to small sample size, missing info, and transformed expression values. GSE50084 has been discarded



```{r}
# Contains "GSE129166", "GSE34748"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[2]])["Gene Symbol"])[[1]])
```

```{r}
# Contains "GSE129166", "GSE46474"
length(union(fData(geo_list[[1]])["Gene Symbol"], fData(geo_list[[5]])["Gene Symbol"])[[1]])
```

All 3 have perfect overlap so the datasets recommended for use are "GSE129166", "GSE34748", and "GSE46474"

(Mukund): Also just thinking, but we probably do want to go through the effort of trying to clean another dataset to match these. firstly for the sake of increased difficulty in the eyes of Jean, but also sometimes without going through the process it can be easy to overlook something in the data cleaning aspect. I am just a bit worried because after talking to some other people they have all told me the data cleaning for this is really hard so not sure if we are actually missing something. 

```{r}
GSE34748 = getGEO(filename="Data/GSE34748_series_matrix.txt.gz")
clinical_GSE34748 = pData(GSE34748)

```
Also just doing some checking and some of these don't actually have clinical data

```{r}
GSE46474 = getGEO(filename="Data/GSE46474_series_matrix.txt.gz")
clinical_GSE46474 = pData(GSE46474)

```
This one has clinical data!!
```{r}
GSE129166 = getGEO(filename="Data/GSE129166_series_matrix.txt.gz")
clinical_GSE129166 = pData(GSE129166)

```
Also no clinical data
```{r}
#GSE129166 = getGEO(filename="Data/GSE129166_series_matrix.txt.gz")
#clinical_GSE129166 = pData(GSE129166)

```

## Mukund CPOP analysis 
```{r}
#CPOP data

GSE36059 = getGEO("GSE36059")[[1]]
GSE48581 = getGEO("GSE48581")[[1]]
GSE129166 = getGEO("GSE129166")[[1]]
GSE15296 = getGEO("GSE15296")[[1]]
```


```{r}
## keeping only the 100 most variable genes in my data frame 
exp_GSE36059 = (exprs(GSE36059))
Variance = rowVars(as.matrix(exp_GSE36059))
Variance = as.data.frame(Variance)
exp_GSE36059 = as.data.frame(exp_GSE36059)
exp_GSE36059 = cbind(exp_GSE36059, variance = Variance)
exp_GSE36059 = slice_max(exp_GSE36059, order_by = Variance, n = 2000)
exp_GSE36059 = subset(exp_GSE36059, select = -c(Variance))
row_names_exp_GSE36059 = rownames(exp_GSE36059)


exp_GSE46474 = (exprs(GSE46474))
Variance = rowVars(as.matrix(exp_GSE46474))
Variance = as.data.frame(Variance)
exp_GSE46474 = as.data.frame(exp_GSE46474)
exp_GSE46474 = cbind(exp_GSE46474, variance = Variance)
exp_GSE46474 = slice_max(exp_GSE46474, order_by = Variance, n = 2000)
exp_GSE46474 = subset(exp_GSE46474, select = -c(Variance))
row_names_exp_GSE46474 = rownames(exp_GSE46474)


exp_GSE48581 = (exprs(GSE48581))
Variance = rowVars(as.matrix(exp_GSE48581))
Variance = as.data.frame(Variance)
exp_GSE48581 = as.data.frame(exp_GSE48581)
exp_GSE48581 = cbind(exp_GSE48581, variance = Variance)
exp_GSE48581 = slice_max(exp_GSE48581, order_by = Variance, n = 2000)
exp_GSE48581 = subset(exp_GSE48581, select = -c(Variance))
row_names_exp_GSE48581 = rownames(exp_GSE48581)


intersection = intersect(row_names_exp_GSE36059, row_names_exp_GSE46474)
intersection = intersect(intersection, row_names_exp_GSE48581)

exp_GSE36059 = as.data.frame(t(as.matrix(exp_GSE36059)))
exp_GSE36059 = subset(exp_GSE36059, select = c(intersection))

exp_GSE46474 = as.data.frame(t(as.matrix(exp_GSE46474)))
exp_GSE46474 = subset(exp_GSE46474, select = c(intersection))

exp_GSE48581 = as.data.frame(t(as.matrix(exp_GSE48581)))
exp_GSE48581 = subset(exp_GSE48581, select = c(intersection))

GSE36059_id <- data.frame("Dataset" = rep("GSE36059",nrow(exp_GSE36059)))
GSE46474_id <- data.frame("Dataset" = rep("GSE46474",nrow(exp_GSE46474)))
GSE48581_id <- data.frame("Dataset" = rep("GSE48581",nrow(exp_GSE48581)))

z1 = exp_GSE36059 %>% as.matrix()
z3 = exp_GSE46474 %>% as.matrix()
z2 = exp_GSE48581 %>% as.matrix()

## arcsine transformation

exp_GSE36059_arc <- exp_GSE36059
exp_GSE36059_arc = exp_GSE36059_arc / max(exp_GSE36059_arc)
exp_GSE36059_arc = asin(sqrt(exp_GSE36059_arc))

exp_GSE46474_arc <- exp_GSE46474
exp_GSE46474_arc = exp_GSE46474_arc / max(exp_GSE46474_arc)
exp_GSE46474_arc = asin(sqrt(exp_GSE46474_arc))

exp_GSE48581_arc <- exp_GSE48581
exp_GSE48581_arc = exp_GSE48581_arc / max(exp_GSE48581_arc)
exp_GSE48581_arc = asin(sqrt(exp_GSE48581_arc))

z1_pairwise = pairwise_col_diff(z1) %>% as.matrix()
z2_pairwise = pairwise_col_diff(z2) %>% as.matrix()
z3_pairwise = pairwise_col_diff(z3) %>% as.matrix()


z1_arc = pairwise_col_diff(exp_GSE36059_arc) %>% as.matrix()
z3_arc = pairwise_col_diff(exp_GSE46474_arc) %>% as.matrix()
z2_arc = pairwise_col_diff(exp_GSE48581_arc) %>% as.matrix()

## log transform

exp_GSE36059_log <- exp_GSE36059
exp_GSE36059_log = exp_GSE36059_log + 1
exp_GSE36059_log = log(exp_GSE36059_log)

exp_GSE46474_log <- exp_GSE46474
exp_GSE46474_log = exp_GSE46474_log + 1
exp_GSE46474_log = log(exp_GSE46474_log)

exp_GSE48581_log <- exp_GSE48581
exp_GSE48581_log = exp_GSE48581_log + 1
exp_GSE48581_log = log(exp_GSE48581_log)

z1_log = pairwise_col_diff(exp_GSE36059_log) %>% as.matrix()
z3_log = pairwise_col_diff(exp_GSE46474_log) %>% as.matrix()
z2_log = pairwise_col_diff(exp_GSE48581_log) %>% as.matrix()


```

## pre transformation plot
```{r}
box11 = cbind(boxplot_tbl(z1, index = 1), GSE36059_id)
box22 = cbind(boxplot_tbl(z3, index = 1), GSE46474_id)
box33 = cbind(boxplot_tbl(z2, index = 1), GSE48581_id)
box4 = rbind(box11, box22, box33)

expressionplot <-
ggplot(data = box4, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  theme(axis.title.x=element_blank()) +
  theme(axis.title.y=element_blank()) +
  ylim(0,15) + 
  theme(legend.position="bottom") +
  theme(legend.title = element_blank()) +
  labs(title = "Raw Data") +
  theme(plot.title = element_text(size=10))

expressionplot
```

## Boxplot to visualise if the arc transformations were good
```{r}
box1_arc = cbind(boxplot_tbl(z1_arc, index = 1), GSE36059_id)
box2_arc = cbind(boxplot_tbl(z3_arc, index = 1), GSE46474_id)
box3_arc = cbind(boxplot_tbl(z2_arc, index = 1), GSE48581_id)
box4_arc = rbind(box1_arc, box2_arc, box3_arc)

arcplot <-
ggplot(data = box4_arc, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Arcsine transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

arcplot
```

## boxplot to see if the log transformation was good
```{r}
box1_log = cbind(boxplot_tbl(z1_log, index = 1), GSE36059_id)
box2_log = cbind(boxplot_tbl(z3_log, index = 1), GSE46474_id)
box3_log = cbind(boxplot_tbl(z2_log, index = 1), GSE48581_id)
box4_log = rbind(box1_log, box2_log, box3_log)

logplot <-
ggplot(data = box4_log, aes(x = object, y = means)) +
  geom_point(aes(color = Dataset), size = 0.1) +
  geom_errorbar(aes(ymin = q1,
                    ymax = q3,
                    color = Dataset), size = 0.1,  alpha = 0.2) +
  ggsci::scale_color_d3() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank()) +
  xlab("Samples") +
  theme(axis.title.y=element_blank()) +
  labs(title = "Log transformation + pairwise difference") +
  theme(plot.title = element_text(size=10))

logplot
```
## getting the results vectors 
```{r}
p_GSE46474 = pData(GSE46474)
p_GSE48581 = pData(GSE48581) 
p_GSE36059 = pData(GSE36059) 

p_GSE36059$diagnosis = ifelse(p_GSE36059$characteristics_ch1 == "diagnosis: non-rejecting", 0, 1)
p_GSE48581$diagnosis = ifelse(p_GSE48581$characteristics_ch1.1 == "diagnosis (tcmr, abmr, mixed, non-rejecting, nephrectomy): non-rejecting", 0, 1)
p_GSE46474$diagnosis = ifelse(p_GSE46474$characteristics_ch1.5 == "procedure status: post-transplant non-rejection (NR)", 0, 1)

y1 = as.factor(p_GSE36059$diagnosis)
y2 = as.factor(p_GSE48581$diagnosis)
y3 = as.factor(p_GSE46474$diagnosis)
### GSE36059
### GSE48581
# these have reject + stable but categorized in more detail -> either have more groups that we are predicting, or we could do purely binary 
```


## running CPOP
```{r}
#CPOP model


counter = 0

bal_acc_y3_test = numeric(length(101))
features_test = numeric(length(101))
y3_alpha_accuracy_list = list()
features_list = list()

for (i in c(71)) {
  set.seed(3888)
  counter = counter + 1
  #alpha_vector = seq(0.01, 1, by=0.01)
  alpha_vector = c(0, 0.01,0.05, 0.5, 0.9, 1)
  counter2 = 0
  for(a in alpha_vector){
    counter2 = counter2 + 1
      cpopresult=
      cpop_model(
        z1_log,
        z2_log,
        y1,
        y2,
        w = NULL,
        n_features = 30,
        n_iter = 20,
        alpha = 1,
        family = "binomial",
        s = "lambda.min",
        cpop2_break = TRUE,
        cpop2_type = "sign",
        cpop2_mag = 1,
        cpop1_method = "normal",
        intercept = FALSE)

    prediction_z3 <- predict_cpop(cpopresult, z3_log, s = "lambda.min")
    c = is.na(prediction_z3$cpop_model_avg_class)
    na_index1 = which(c %in% c(TRUE))
    
    if (length(na_index1) == 0) {
      bal_acc <- yardstick::bal_accuracy_vec(
      truth = y3,
      estimate = factor(coalesce(prediction_z3$cpop_model_avg_class, "0")))
    }
    else {
      bal_acc <- yardstick::bal_accuracy_vec(
      truth = y3,
      estimate = factor(coalesce(prediction_z3$cpop_model_avg_class, "0")))
    }
    
    bal_acc_y3_test[counter2] = bal_acc
    features_test[counter2] = length(cpopresult$features)
  }
  
  y3_alpha_accuracy_list[[counter]] = bal_acc_y3_test
  features_list[[counter]] = features_test
  print(y3_alpha_accuracy_list)
}


```

